Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia

### 2020-09-09 internal defense feedback/Answers-jury-members-report.tex
One before the very last

	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.

@@

\begin{itemize}
	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider? Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?

@@

	\item  What recommendation up to scale/order can method work with. What are the limitations? It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...? A more detailed discussion is desirable.
	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).

@@

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

@@

    \renewcommand{\thefigure}{6.9}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. \color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively. \color{black}  }
    \end{figure}

@@

	\color{blue}
	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
	The damped Newton method  
	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
	The continuous Newton method is a sequence of damped Newton methods. To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.

@@

    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$. The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$. Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
    Thus,  
	\color{black}

	\item  About trying to find a good transformation from LS to EIV model. I am wondering if you can think of other more accurate starting points instead of LS. Closer to the problem you want to study, but still preserving computational lightness? The closer you start from objective, better the Taylor expansion would be.. Something to investigate as well...
	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,... Perhaps the use of median of means plugged into your approach might produce interesting alternatives for outliers or heavy tailed data. But this is an intricate thing to deal with. Might be interesting to ponder this in future.
	\item  On deep learning: could this be an alternative?
\end{itemize}

@@

	\item  Chapter 2, Section 2.2.1. Can you give me some classic solutions/algorithms instead of LS in general? (Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse. Do you know another solution where they don't use K on the left-hand-side but something else? To get rid of noise by using a matrix on the left-hand-side, is related to EIV. There are some recursive solutions for this. It could be interesting to compare with such IV solutions.
	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
	\item  Equation 3.1: an LTI system, described with SS representation. You only put output noise on this representation. People sometimes also add process noise. Why? How to modify your approach if there is process noise? Is it still white in this case?
	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations! there is no epsilon (noise) in the equation. But then you do an SVD of H and estimate from there. How does it work if you have noise? Have you tried implementing it and see what you get?

@@

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.

@@

	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ... (even if randomly generated). You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?

@@

	{\bfseries \color{red} Interesting question, I want to try it this evening \color{black}}

@@

	\item   On the impact of selected order on results. You were talking about a related bias/variance trade off. So, do you know in the literature some solutions to avoid overfitting? Can you cite something? Can you use something in your case to test for a good order of system?
	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation? Have you tried this? Can this be used here to select good orders? Better results? Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set. with this i should see what is the best value... Have you tried this?
	\item   In Section 6.2.2. Why do you call it a max likelihood approach? To me it is not one. Why do you call it like this? Where is the likelihood? Depending on which likelihood (eg. Gaussian, ...) it leads to LS. Don't' call it ML if you dont give distributions on ... In this case it is better to call it a nonlinear minimization procedure. Not clear why you call it ML.

@@

	\item   Concerning the ML approach in simulations, you say that it takes 30 seconds to complete. What is done exactly in these 30 seconds; how many samples, realizations,... I am a bit surprised you require 30 sec for 2 parameters.
\end{itemize}

@@

\begin{itemize}
	\item When solving algorithms. Do you opt for closed-form solutions, or do you use optimization based algorithms? How to solve LS problems? Do you solve them using pseudo-inverse or do you do gradient descent?
	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly? This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
	\item  What if the $K^\top K$ matrix is not invertible? What if you have fewer samples than unknowns; what can you do? Solutions that assume structure like in compressed sensing, regularization,... can you use them, what would be the impact of that?
	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
	\item  You mentioned deep learning could be applied; could you show how? An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR. This clogs performance; any ideas? Is this a limitation?
	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".

>>>>

	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).

	{\bfseries In Section 3.2 was added a discussion on the scalability of the data-driven step input estimation method.}

	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.

@@

\begin{itemize}
	\item You develop results for uncertainty quantification; what type of uncertainties can your approach deal with and which levels of noise can you consider? Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.

	{\bfseries A numerical sensitivity analysis was conducted on both subspace and maximum-likelihood methods for solving the affine input estimation problem. The sensitivity analysis showed that the subspace method can handle uncertainty on the generating system parameters up to 5\% without afecting deeply the performance of the method. Moreover, the mean-squared error of the estimation computed by the subspace method is similar in up to one order of magnitude in the SNR interval from 20 to 60 dB, being $10^{-1} \ \mathrm{g}^2$ its maximum value at 40 dB. The levels of measurement noise in practical applications are found in this SNR interval. }

	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?

@@

	\item  What recommendation up to scale/order can method work with. What are the limitations? It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...? A more detailed discussion is desirable.

	{\bfseries In Section 3.2 was added a discussion on the scalability of the data-driven step input estimation method. Since it uses recursive least squares to update the estimation, its complexity is linear in the number of parameters $n+1$. This means that for small number of samples, $N=100$, or large number os samples, $N=10^4$, the complexity per recursion does not change.

	The computation of the bias and variance of the input estimation, as it is proposed in Chapter 4, has a complexity $O(n^3N)$. This complexity prevents the implementation in real time of these statistical moments estimation. A future research work can be devoted to find efficient estimation of these input estimation moments, to be obtained right after the input estimation itself. }

	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).

@@

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

@@

    \renewcommand{\thefigure}{6.9}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../ChapterRampInput/fig/Fig_8.pdf} 
    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. \color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively. \color{black}  }
    \end{figure}

@@

	{\bfseries The investigation of the error bounds for the second order Taylor expansion can be conducted using the continuous Newton method. This is proposed as future research work in Chapter 7.}

	\color{blue}
	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
	The damped Newton method  
	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
	The continuous Newton method is a sequence of damped Newton methods. To implement the continuous Newton method, select $0<T<1$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.

@@

    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$. The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$. Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
    The questions that remain open are how large can $r$ be?, and how is the radius $r$ related to the perturbation noise variance? 
	\color{black}

	\item  About trying to find a good transformation from LS to EIV model. I am wondering if you can think of other more accurate starting points instead of LS. Closer to the problem you want to study, but still preserving computational lightness? The closer you start from objective, better the Taylor expansion would be.. Something to investigate as well...

	{\bfseries To get starting points to the EIV model, that are more accurate than LS, we would need to incorporate more prior knowledge. In the formulation of the EIV problem we have made assumptions of the model class for the sensor and the Gaussian white noise of given variance for the additive measurement noise. The LS is an heuristic algorithm that was chosen by its simplicity for solving online the estimation problem by means of recursion. However, the LS solution is biased. The optimal solution could be obtained with a maximum likelihood algorithm provided we have a large amount of samples, but we need to provide approximations starting from small number of samples. Moreover, the maximum likelihood solution of the EIV problem needs structured total least squares, which does not have a closed form that could be extended with a Taylor series, or structured total maximum likelihood, which do not support the correlation between the regression matrix and the regressor in the EIV problem. The recursive versions of these two structured approaches is also more computational expensive than RLS. Therefore, it is required to continue the research towards efficient solution algorithms.  }

	\item  About robustness: is it an issue in practical problems with heavy tailed noise, outliers,... Perhaps the use of median of means plugged into your approach might produce interesting alternatives for outliers or heavy tailed data. But this is an intricate thing to deal with. Might be interesting to ponder this in future.

	{\bfseries There are some estimators that can deal with heavy tailed noise. In the paper \cite{Balkema18} there  are described different estimators for a linear regression \linebreak $Y_i = b+a X_i + Y_i^{*}, \quad i = 1,\ldots,n$, where both the explanatory variables $X_i$ and the errors $Y_i^{*}$ have heavy tails. The $X_i$ are samples from a Pareto distribution on $(1, \infty)$ arranged in decreasing order $X_n<\cdots<X_2<X_1$, as if they were generated from the order statistics $U_1<\cdots<U_n$ of a sample of uniform variables in $(0,1)$ by setting $X_i = 1/U_i^\zeta$. The parameter $\zeta$ is the tail index of the Pareto distribution. For the variables $Y_i^*$, the tail index is $\eta$. 
	Heavy tails are indicated by larger tail indexes. For $\eta < 1/2$, the variable $Y^*$ has finite variance and least squares is the suggested estimator. For $\zeta < 1/2$, the variable $X$ has finite variance and the least absolute deviation is the suggested estimator, otherwise the Theil-Sen estimator is preferable. The paper presents empirical results comparing the estimation bias and standard deviation for different tail indexes.

	Regarding outliers, in \cite{Maronna06} is described an approach of using least squares to identify the outliers, followed by a recalculation without considering the outliers. In this sense, one could also consider that the removed outliers are missing data, and use low-rank approximation. The classical approach is using a maximum likelihood type estimator (M-estimators) as it is described in \cite{Zhang97} and \cite{Huber81}. The M-estimators aim to mitigate the outliers impact by using a function of the residual $\rho(r_i)$ instead of the the squared residuals $r_i^2$ in the minimization $\text{min} \ \sum{\rho(r_i)}$. The function $\rho$ is symmetric, positive definite, with only one minimum at 0, and increasing slowly than the the square.

	Selecting another estimator to solve the EIV problem constructed by the data-driven step input estimation problem to deal with heavy tailed noise, or with outliers, might have a big impact on the implementation. The implementation may be computationally complex and not suitable for online applications.

	}

	\item  On deep learning: could this be an alternative?

	{\bfseries Deep learning could be used but it would be an overkill methodology for linear estimation. However, for metrology applications were the input is the to-be-estimated signal, implementing deep learning could be a challenge. The neural networks are suitable for non-linear model based applications were the training require the design and application of input stimuli to correlate with the corresponding outputs of the system. 
	In metrology, it is not possible to train deep learning methodologies with real input data, because the input data is unknown and cannot be observed. 
	By applying neural networks trained by simulation would estimate inaccurately the desired inputs.  }

@@

	\item  Chapter 2, Section 2.2.1. Can you give me some classic solutions/algorithms instead of LS in general? (Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse. Do you know another solution where they don't use K on the left-hand-side but something else? To get rid of noise by using a matrix on the left-hand-side, is related to EIV. There are some recursive solutions for this. It could be interesting to compare with such IV solutions.

	{\bfseries In Section 2.2.1 the follwing paragraphs have been modified to include references of instrumental variable solutions to the errors-in-variables problems.}

	\color{blue}
	Instead of using a total least-squares (TLS), or an intrumental variables, solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
    In the review conducted by \cite{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
    the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed. 
    The classical TLS solution is obtained using singular value decomposition of the augmented matrix built from the regression matrix and the regressor as additional columns.
    However, the studies presented in \cite{VanHuffel07TLSeditorial} demonstrate that the structured TLS solutions cannot be generalized since each structure in an EIV problem requires a specific treatment. 
    Moreover, the ML estimator of structured EIV problems leads to non-convex optimization, and the global optimum is not guaranteed \cite{Beck10}.
    On the other hand, the instrumental variable (IV) solution of EIV problems use an additional variable.
    The additional variable is an instrument with the purpose of improving the consistency of the estimate.
    To do this, the construction of the instrumental variable has to be done in such a way that it is uncorrelated with the perturbations and correlated with the regressor, as it is described in \cite{Soderstrom18}. 
    Unfortunately, the computational complexity of the TLS, ML, and IV solutions inhibits their real-time implementation.

    The solution of the step input estimation problem is preferred to be online, since its applications are in the field of metrology.
    In the literature there are references for online implementations of TLS, ML or IV methods, which are are mainly recursive versions of the solution methods.
    These recursive versions are developed to solve EIV problems for system identification and fault detection.
    We have, for example, generalized recursive TLS solutions introduced in \cite{Rhode14recursive} and \cite{Rhode14recursivecov}, and recursive IV solutions developed in \cite{Djouambi12}, \cite{Shang16} and \cite{Gil15}.
    These methods cannot be used as off-the-shelf solutions to the step input estimation EIV problem because they require to be adapted to the specific requirements of the problem.
    One is that the sensor output is the only measured signal, since the input is to-be estimated, and the solution should be simpler, preferable with a linear complexity $O(n)$, instead of the cubic $O(n^3)$ or cuadratic $O(n^2)$ complexities of the mentioned recursive methods, where $n$ is the sensor order.
    The proposal of the data-driven step input estimation method includes the recursive least-squares (RLS) solution.
    Compared to recursive TLS or recursive IV solutions, RLS is suboptimal but more simple solution to the structured EIV problem with lower computational complexity,  suitable for real-time implementation.
	\color{black}

	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?

	{\bfseries The first paragraph of Chapter 3 has been added to explain this concept:  }

	\color{blue}
    The dynamic systems in the real world are non linear. 
    Under certain intervals of operation, the linear approximations can explain the dynamics of the systems. 
    The linear and time invariant systems have been studied in depth and the linear theory is a solid collection of tools to analyze and explain the behaviour of dynamic systems in a wide range of applications. 
    On the other hand, the study of non-linear systems is more complex and strongly depends on the type of non-linearity that each particular systems exhibit. 
    When the non-linear modeling does not apport a big amount of benefits into the application study, such as significative accuracy improvement or supporting wide ranges of operation, then is better to resort to the utilization of LTI representations. 
    A sensor is a dynamic system whose operation range is preferred to be linear, to facilitate the repeatability and reproducibility of the measurements.
    This means that every time the sensor must deliver identical responses for the same input excitations.   
    \color{black}

	\item  Equation 3.1: an LTI system, described with SS representation. You only put output noise on this representation. People sometimes also add process noise. Why? How to modify your approach if there is process noise? Is it still white in this case?

	{\bfseries In Chapter 3 was added the following paragraph to explain the impact of white process noise in the response of al LTI system. }

	\color{blue} 
    The process noise is the representation of the uncertainty in the system state due to unmodeled perturbations that affect the state evolution. 
    The process noise is modeled as an additional input to the system excitation. 
    For a linear time invariant (LTI) system, the response to a sum of two input excitations is the sum of the two responses that correspond to the individual inputs.
    It is demonstrated in \cite{Gubner06} and \cite{Smith11} that the response of an LTI system to a white noise input is colored noise, since the power spectral density of the response is not constant, but instead proportional to the squared magnitude of the system impulse response.
    Nevertheless, in the bandwidth operation of a properly designed measurement system, processing white noise results in approximately white noise output. 
    Moreover, in the present work the measurement noise exists only in the system response signal, because the input is not direcly observed, and the sensor operation is not subject to external perturbation. 
    Under these conditions, the mean and variance of the measurement noise $\epsilon$ describe implicitly the process noise entering the system. 
    \color{black}

	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations! there is no epsilon (noise) in the equation. But then you do an SVD of H and estimate from there. How does it work if you have noise? Have you tried implementing it and see what you get?

@@

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.6) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance.
    \color{black}

@@

	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ... (even if randomly generated). You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.

	{\bfseries Section 4.2 has been improved to explain how the exact data and the perturbation noise are generated. This modification allows to reproduce the simulations conducted and understand the outcome of them. }

	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?

@@

	{\bfseries The numerical study of the expressions that calculate the step input estimation bias and variance has been appended to include the results of using unstructured perturbation in the expressions for the structured case, and also structured perturbation in the expressions for the unstructured case. It is expected that the performance is very far from that obtained when using the expressions in concordance with the type data they were derived for. A discussion on this observations is included in Section 4.2.3.}

@@

	\item   On the impact of selected order on results. You were talking about a related bias/variance trade off. So, do you know in the literature some solutions to avoid overfitting? Can you cite something? Can you use something in your case to test for a good order of system?

	{\bfseries If we evaluate the step input estimation bias and variance obtained by the data-driven method for different model orders, we expect the results follow the bias-variance tradeoff. It is commonly accepted that for decreasing orders the bias is large since not all dynamics are captured, and that for increasing orders the bias is small and the variance is large because the system dynamics are captured and also the measurement noise is being modeled. However, after observing the results in the experimental validation of the method, we get better performance by overfitting. This could be valid only for some estimators. 
	For example, in the research work of \cite{Neal19} it is described a neural network that does not follow the bias variance tradeoff, contradicting \cite{Geman92} and \cite{Manning08}, who state that an overparameterized neural network reduces the bias by interpolating the data with a significant variance contribution to the mean-squared error.

	Some approaches described in the literature to avoid overfitting include crossvalidation (CV), minimum description length (MDL) and the Akaike information criterion (AIC).
	These methodologies are described in books such as \cite{Hastie09}  and \cite{James13}. 
	In the paper \cite{Arlot10}, CV and MDL are compared trying to answer whether penalized criteria in classical AIC should be preferred to CV for a given model selection problem. 
	On one hand, provided the data is i.i.d., CV gives good model selection in almost any framework. 
	On the other hand, CV can be less accurate, even with larger computational cost, than procedures designed to be optimal in a specific framework, such as AIC.

	To describe the classical methods, the following paragraph has been added in subsection 5.2.
	}

	There exist methods for estimating the performance of different models that are useful to choose among the models.
	The model selection methods can be analytical, such as the Akaike information criterion (AIC) introduced in the paper of \cite{Akaike74}, the minimum description length (MDL) iniciated by the work of \cite{Rissanen78}, or numerical such as crossvalidation.
	The AIC is defined for a class of models fit by maximum likelihood. 
	For the linear regression problem with Gaussian perturbation (3.20), the AIC is given by
	\begin{equation} \tag{5.9} \mathrm{AIC} = 2 \left( n+1 \right) - \mathrm{ln} \left( \widehat{L} \right) \label{eqn:AIC} \end{equation}
	where $n+1$ is the number of estimated parameters, and $\widehat{L}$ is the maximum value of the likelihood function of the model.
	The AIC rewards the goodness of fit by considering the maximum likelihood value, and adds a penalty term that increases the cost with respect to the number of parameters to prevent overfitting.
	In the opinion of \cite{McElreath20}, the AIC is reliable when the priors overwhelmed by the likelihood, the posterior distribution is approximately multivariate Gaussian, and the sample size is much larger than the number of parameters.
	Using this criterion for model selection is simple, \cite{Hastie09} suggest just picking the model that provides the smallest AIC value.

	The MDL approach is motivated from an optimal coding viewpoint, aiming to choose the most parsimonious model, that, according to \cite{Hastie09}, is the shortest model that can encode the data for transmision of messages.
	If we have a model with parameters $\theta$, and data $(K, y)$, with the conditional probability of the outputs $\mathrm{P}(y | \theta, K)$, and assuming the receiver knows the inputs and we transmit the outputs, then the message length required to transmit the outputs is
    \begin{equation} \tag{5.10} \mathrm{length} = - \mathrm{ln} \left( \mathrm{P}( \theta ) \right) - \mathrm{ln} \left( \mathrm{P}(y | \theta, K) \right) . \label{eqn:MDL} \end{equation}
    The length is the sum of the code length required for transmitting the model parameters with the code length required for transmitting the discrepancy between the model and the actual target values. 
    The recommended model is the one that minimizes (\ref{eqn:MDL}).
	\cite{McElreath20} arguments that model fitting is a form of data compression, where parameters explain the relationships among the data. 
	The parameters compress the data into a simpler form, with some loss of information about the sample, and 
	can be used to generate new data by decompression.
	Learning about the data requires using a simpler model that achieves some compression, but not too much.
    \color{black}

	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation? Have you tried this? Can this be used here to select good orders? Better results? Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set. with this i should see what is the best value... Have you tried this?

	{\bfseries The crossvalidation method has not been applied on the experimental data. This data contains the step responses of a weighing system. The step responses are acquired with a sampling frequency of 4 kHz. One difficulty for applying crossvalidation is the division of the data. If we select the first 2/3 of the step response samples for order estimation and the last 1/3 for validation, we may leave the data with the diminished transient efcects in the validation set. Another approach could be taking every second sample for validation, dividing the measured responses in two by downsampling. However, this approach requires filtering against possible aliasing because the frequency characteristics of the response signal change. A third way that could be tried is dividing the 100 measured responses in two sets, devoting 80 responses for order estimation and 20 responses for validation.   }

	\item   In Section 6.2.2. Why do you call it a max likelihood approach? To me it is not one. Why do you call it like this? Where is the likelihood? Depending on which likelihood (eg. Gaussian, ...) it leads to LS. Don't call it ML if you dont give distributions on ... In this case it is better to call it a nonlinear minimization procedure. Not clear why you call it ML.

@@

	\item   Concerning the ML approach in simulations, you say that it takes 30 seconds to complete. What is done exactly in these 30 seconds; how many samples, realizations,... I am a bit surprised you require 30 sec for 2 parameters.

	{\bfseries The large time consumption in the local optimization iterations is due to the different functions that take place during the minimization of the error function. The iterations are performed in a receding horizon scheme, starting from a small amount of samples (50) and repeating the error minimization for each new subset of samples acquired (5), until covering all samples (1000). In each iteration, the response of a dynamic system is simulated several times, for the different tried optimization variables. This simulated response is compared to the actual measured response, aiming to reduce the error between these two responses. Every time that a sensor response is simulated for a particular vector of optimization variables being tested, the analytic Jacobian is also evaluated. The evaluation of the Jacobian matrix requires the  simulation of one additional dynamic system per to-be-optimized parameter. Considering that the simulations of all these responses take considerable amount of time, the elapsed time that the ML method takes is 25 s in average. Then, there is a need for efficient methods that could run in real time to optimize online the input model parameters and the sensor model parameters of interest. }

@@

\begin{itemize}
	\item When solving algorithms. Do you opt for closed-form solutions, or do you use optimization based algorithms? How to solve LS problems? Do you solve them using pseudo-inverse or do you do gradient descent?

	{\bfseries The closed form solutions allow us to describe the operations in a more elegant way. The closed form solutions implies batch processing of the operations. The formulation of the problem also gives a hint of which solution to choose. For instance, an unconstrained linear minimization problem can be solved with least squares but constrained nonlinear minimization problems most likely require optimization. There are different methods to solve least squares problems. The pseudo-inverse approach is the first option that comes to my mind, but modern numerical software, such as Matlab and julia, have toolboxes that solve efficiently the least squares problems depending on the regression matrix and regressor data. Regarding the maximum-likelihood method of Chapter 6, the optimization method uses gradient descend to find the direction towards the local minimum. Other minimization problems in low-rank approximation can be solved using variable-projection optimization methods. }

	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly? This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?

	{\bfseries The increase in dimensionality is explained by the current trend in the industry of installing more sensors. More signals are used to monitor the production lines. The transference of data is relevant for decision making. The data-driven input estimation method can be used to estimate recursively the input applied to the sensors, and to tract the evolution of inputs. In this way, the estimation method facilitates the controllers by providing online estimations of the inputs. This improves the performance of closed form solutions, which need more computational resources and time more time for the batch processing of data.}

	\item  What if the $K^\top K$ matrix is not invertible? What if you have fewer samples than unknowns; what can you do? Solutions that assume structure like in compressed sensing, regularization,... can you use them, what would be the impact of that?

	{\bfseries A possible lack of invertibility of the matrix $K^\top K$ may be due to missing data. A different methodology to low-rank approximation is compressed sensing, where the missing data is interpreted as sparsity. Under the assumption that a signals is sparse because is subject to many near zero coefficients in a given basis, then changing to another basis and taking a weighted linear linear combination of the available samples would lead to the estimation of the missing data values. These idea is described as future work in the final chapter of the thesis  }

	\subsection*{Compressed sensing methods for recovering missing data}

	Compressed sensing is a methodology that is being developed quite fast and that can offer an alternative to perform measurements with missing data. 
	The aim of compressed sensing is the recovery of the missing samples of a signal by computing a sparse coefficient vector, under the assumption that the signal belongs to a generating system of larger order. 
	The missing data estimation problem can be expressed as an overdetermined system of equations with a sparse regression matrix. 
	The review presented in \cite{Eldar2012compressed} provides an introduction of the mathematical modeling of compressed sensing, and gives the conditions for sparse recovery. 
	The recovery algorithms include convex optimization, greedy and combinatorial algorithms, to name a few.

    According to \cite{Rani18}, compressed sensing is applied when we canot afford buying enough sensors like in non-visible wavelengths applications, when the measurements themselves are expensive as in neutron-scattering imaging, when measurements take a lot of time as in medical imaging, and when a constrain of the measurement is the power consumption.
    A particular field of application for compressed sensing is the recovery of bioelectrical signals, as it is detailed in \cite{Craven15}.

    Another good review of compressed sensing is found in \cite{Duarte11}, where it is explained that the  research community is splitted into a group that studies the problems in which the compressed sensing matrices exhibit structure due to the measurement system modeling, and another group which considers that the signal structured representations not only depend on sparsity but instead assume that the signals belong to continuous time and have infinite-dimensional representations. The solutions provided exploit these structures in the development of the algorithms. Some examples of these structure matrices are described in \cite{Kezhi15} including Gaussian, Toeplitz, Hadamard, and structurally random matrices.

    It might be interesting to compare the solutions of the low-rank approximation methods with compressed sensing in terms of computational complexity and feasibility for online implementations. 
    Therefore, it is proposed as a future research to develop a compressed sensing algorithm, that can estimate directly the input true value by processing the sensor response to the unknown input, and in presence of missing data. 
	\color{black}

	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?

	{\bfseries A discussion on the scalability of the data-driven step input estimation method has been added in Section 3.2. The method uses recursive least squares to update the input estimation value, thus its complexity is linear in the number of parameters $n+1$, and independent from the number of samples $N$.

	The computation of the bias and variance of the input estimation, which is described in Chapter 4, has a complexity $O(n^3N)$. This complexity prevents the online implementation of the estimation bias and variance. It is proposed as future research work to study efficient methods to estimate these statistical moments. The objective of this future work is to estimate the input and its mean squared error in real time. }
